Star-galaxy Classification Using Deep Convolutional Neural Networks

### 2016-08-15 paper/main.tex
remove sout and red text

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the reduced, calibrated pixel values that comprise an image into suitable features,
such as magnitudes or shape information of an object.
\changed{
In a branch of machine learning called \textit{deep learning}~\citep{lecun2015deep},
features are not designed by human experts;
they are learned directly from data by deep neural networks.
Deep learning methods learn multiple levels of features by
transforming the feature at one level into a more abstract feature at a higher
level.
For example, when an array of pixel values is used as input to a deep learning
method, the features in the first layer might represent locations and orientations
of edges.
The second layer could assemble particular arrangements of edges into more
complex shapes, and subsequent layers would detect objects as
combinations of low-level features.
These multiple layers of abstraction progressively amplify aspects of the input
that are important for classification tasks.
}
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has not yet been applied to the problem of source classification.

@@

\changed{
To demonstrate the performance of our ConvNet model, we use photometric and
spectroscopic data sets with different characteristics and compositions.
In this section, we briefly describe these data sets and the image pre-processing
steps for retrieving cutout images.
}

@@

Using the CasJobs server, we randomly select a total of 65,000 sources, which are either stars or galaxies.
We exclude some bad photometric observations as follows.
We consider only objects
\sout{ whose $r$-band magnitude is less than 40; there are }
with no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the 65,000 objects we have selected,
and to center each object in the cutout image.
\changed{
Magnitudes in the SDSS photometric catalog are expressed as inverse hyperbolic
sine magnitudes~\citep[also known as luptitudes;][]{lupton1999modified},
and we follow the SDSS convention and convert all flux values to luptitudes.
}
Finally, in order to account for the effect of Galactic dust,
extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have 17,344 stars and 47,656 galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
\sout{ The $ugriz$ color-color space of the sample is plotted in Figure ?.}
We randomly split the objects into training, held-out validation,
and blind test sets of size 40,000, 10,000, and 15,000, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.
The first two panels of Figure~\ref{fig:sdss_mag} show the number of objects
and the fraction of stars in the test set as functions of $r$-band magnitude.
Similarly, Figure~\ref{fig:sdss_g_r} shows the number of objects and the
fraction of stars in the test set as functions of $g-r$ color.
The normalized kernel density estimate distributions for the training
and validation sets are almost identical to those of the test set,
and they are nearly indistinguishable when overlapped.
We do not show the distributions for the training and validation sets in
Figures~\ref{fig:sdss_mag} and \ref{fig:sdss_g_r} to avoid cluttering the
plots.

@@

In the end, we have 8,545 stars and 57,843 galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
\sout{ The $ugriz$ color-color space of the sample is plotted in Figure ?.}
We randomly split the objects into training, held-out validation, and
blind test sets of size 40,000, 10,000, and 13,278, respectively.
Figures~\ref{fig:clens_mag} and \ref{fig:clens_g_r} show the distribution
of objects in the test set as functions of $i$-band magnitude and $g-r$ color.
We do not show the distributions for the training and validation sets,
since the normalized kernel density estimate distributions for the training
and validation sets are almost identical to those of the test set.

@@

\changed{
Neural networks have many hyperparameters, including those that specify the
network itself (\eg the size and non-linearity of each layer)
and those that specify how the network is trained
(\eg the mini-batch size or the learning rate).
Furthermore, the architecture of a neural network can have a
significant impact on its performance.
In this section, we provide a brief description of key hyperparameters in our
ConvNet model, and also present the network architecture.
}

@@

\sout{ A neuron in the human brain receives signals from other neurons through synaptic connections.  If the combination of these signals exceeds a certain threshold, the neuron will fire and send a signal to other neurons.  Intelligence is believed to be the collective effect of approximately $10^{11}$ neurons firing.  }
An artificial neuron in most artificial neural networks is represented
as a mathematical function that models a biological neural structure
\changed{
\citep{aggarwal2014data}.
A schematic representation is shown in Figure~\ref{fig:neuron_a}.
}
Let $\bmath{x}=\left(x_1,x_2,\dots,x_n\right)$ be a vector of inputs to a given neuron,
$\bmath{w}=\left(w_1,w_2,\dots,w_n\right)$ be a vector of weights, and
$b$ be the bias.
Then, the output of the neuron is
\changed{
\citep{rosenblatt1961principles}
}
\begin{equation}
  y = \sigma \left( \bmath{w} \cdot \bmath{x} + b \right),
  \label{eq:neuron_output}
\end{equation}
where $\sigma$ is the activation function (or \textit{non-linearity}).
\sout{ The sigmoid function, $\sigma(x)=1/\left(1+e^{-x}\right)$, has frequently been used as an activation function, but it has recently fallen out of favor, and the hyperbolic tangent function, $\sigma(x)=\tanh(x)$, is generally used instead.  }
The most popular non-linearity at present
\sout{, however, }
is the rectified linear unit
\citep[ReLU;][]{nair2010rectified}, $\sigma(x)=\max(0, x)$.
ReLUs generally allow much faster
training of deep neural networks with many layers.
However, ReLU units can sometimes result in dead neurons whose output is always zero.
To mitigate this problem, we use leaky ReLUs
\changed{
~\citep{maas2013rectifier}
}
that have a small, non-zero slope in the negative region,
\begin{equation}
  \sigma(x) =
    \begin{cases}
      x & \mbox{if } x \geq 0 \\
      0.01x & \mbox{if } x < 0.
    \end{cases}
\end{equation}

Many deep learning models use feedforward neural network architectures with multiple layers,
where each neuron in one layer is connected to the neurons of the subsequent
layer~\citep{lecun2015deep}.
A schematic representation is shown in Figure~\ref{fig:neuron_b}.
All layers except the input and output layers are conveniently called hidden layers.
\sout{ With multiple hidden layers, deep neural networks can model complex nonlinear relationships, and they have had great success in multimedia tasks where humans traditionally outperformed computers, such as speech recognition, language processing, and image classification.  }

\changed{
We
}
find a set of weights and biases such that,
given $N$ samples, the output from the network
$\bmath{y}=\left(y_1, y_2, \dots, y_N \right)$ approximates the desired output
$\hat{\bmath{y}}=\left(\hat{y}_1, \hat{y}_2, \dots, \hat{y}_N \right)$
as closely as possible for all input
$\bmath{X}=\left(\bmath{x}_1,\bmath{x}_2,\dots,\bmath{x}_N\right)$.
We can formulate this as the minimization of a loss function
$L(\bmath{y},\hat{\bmath{y}})$
over the training data.
In this work, we use \textit{cross-entropy}
\changed{
\citep[also called log loss;][]{murphy2012machine} as the loss function.
For binary classification, the cross-entropy per sample is given by
\begin{equation}
  L(y_j, \hat{y}_j) = -\hat{y}_j \log_2 y_j - (1 - \hat{y}_j) \log_2(1 - y_j),
  \label{eq:cross_entropy}
\end{equation}
}
where $\hat{y}_j$ is the actual truth value (\eg 0 or 1) of the $j$-th data, and
$y_j$ is the probability prediction made by the model.
We compute the loss function by taking the average of all cross-entropies in the sample.
Thus, the loss function becomes
\begin{equation}
\changed{
  L(\bmath{y}, \hat{\bmath{y}})
}
= - \frac{1}{N} \sum_{j=1}^{N} \hat{y}_j  \log_2 y_j
    + (1 - \hat{y}_j)  \log_2 (1 - \hat y_j).
  \label{eq:cross_entropy_all}
\end{equation}

To find the weights $\bmath{w}$ and biases $\bmath{b}$ which minimize the loss,
we use a technique called \textit{gradient descent},
where we use the following rules to update the parameters in each layer $l$:
\begin{align}
  \bmath{w}_l &\rightarrow
  \bmath{w}_l^{\prime}
  = \bmath{w}_l - \eta \frac{\partial L}{\partial \bmath{w}_l} \nonumber \\
  \bmath{b}_l &\rightarrow
  \bmath{b}_l^{\prime}
  = \bmath{b}_l - \eta \frac{\partial L}{\partial \bmath{b}_l},
  \label{eq:gradient_descent}
\end{align}
where $\eta$ is a small, positive number known as the \textit{learning rate}.
\changed{
The gradients can be computed using the backpropagation
procedure~\citep{rumelhart1988learning}.
A common approach to speed up training is to split the training data
into mini-batches~\citep{lecun1998efficient}.
In mini-batch gradient descent, instead of computing the gradients in
Equation~\ref{eq:gradient_descent} for the entire training data,
we only compute the gradient of randomly chosen training examples at each step.
As training examples are usually correlated,
the gradient computed from each mini-batch is a good approximation
of the overall gradient~\citep{bottou1998online}.
As a result, mini-batch gradient descent results in much faster convergence.
However, there is a trade-off: the lower the batch size is,
the lower the convergence rate will be;
the higher the batch size is, the longer it will take to compute the gradient
at each step~\citep{bousquet2008tradeoffs}.
Thus, a moderate batch size, combined with a decaying learning rate, is
generally used in practice. We use a batch size of 128 in this work.
}

@@

\sout{ The gradients in Equation~\ref{eq:gradient_descent} can be computed using the \textit{backpropagation} procedure, which is nothing more than an application of the chain rule for derivatives.  The derivative of the loss function with respect to the input of a unit can be computed by working backwards from the gradient with respect to the output of the unit.  At the output layer, the derivative of the loss function with respect to the output is given by }

\sout{$ \frac{ \partial L }{ \partial y_i } = y_i - \hat{y}_i $}
\sout{$ \frac{ \partial L }{ \partial x_i } = \frac{ \partial L }{ \partial y_i } \frac{ \partial y_i }{ \partial x_i }.  $}

\sout{ At each hidden layer, the derivative of the loss function with respect to the output of each unit is a weighted sum of the derivatives with respect to the inputs of the subsequent layer, }
\sout{$ \frac{ \partial L }{ \partial y_j } = \sum_{i} w_{ji} \frac{ \partial L }{ \partial x_i} $}
\sout{$ \frac{ \partial L }{ \partial x_j } = \frac{ \partial L }{ \partial y_j } \frac{ \partial y_j }{ \partial x_j }.  $}

\sout{ Finally, once the gradients $\partial L / \partial x_j$ have been computed, it is straightforward to compute the gradients for the weight, }
\sout{$ \frac{\partial L}{\partial w_{jk}} = y_k \frac{\partial L}{\partial x_j}.  $}

@@

The convolutional neural network~\citep[ConvNet;][]{fukushima1980neocognitron,lecun1998gradient}
is a type of deep, feedforward neural network
that has recently become
\changed{
a popular
}
approach in the computer vision community.
In a typical ConvNet, the first few stages are composed of two types of layers: 
convolutional layers and pooling layers.

\changed{
The input to a convolutional layer is an image, and the output channels of
each layer are called \textit{feature maps}.
}
To produce output feature maps, we convolve each feature map with a set of weights
called \textit{filters},
and apply a non-linearity such as ReLU to the weighted sum of these convolutions.
Different feature maps use different sets of filters,
but all neurons in a feature map share the same set of filters.
Mathematically, we replace the dot product in Equation~\ref{eq:neuron_output}
with a sum of convolutions. Thus, the $k$-th feature map is given by
\begin{equation}
  y^k = \sigma \left( \sum_{m} \bmath{w}_m^k \ast \bmath{x}_m + b^k \right),
\end{equation}
where we sum over the set of input feature maps,
$\ast$ is the convolution operator, and $\bmath{w}_m^k$ represent the filters.

\changed{
Typically, a
}
pooling layer computes the maximum of a local $2\times2$ patch in a feature map
\changed{
\citep{krizhevsky2012imagenet}.
}
Since the pooling layer aggregates the activations of neighboring units from the previous layer,
it reduces the dimensionality of the feature maps and
makes the model invariant to small shifts and distortions
\changed{
\citep{boureau2010theoretical}
}
.
Two or more layers of convolution and pooling are stacked,
followed by more convolutional and fully-connected layers.

@@

The architecture of \cite{krizhevsky2012imagenet}
uses relatively large receptive fields ($11\times11$) in the first convolutional layers.
\citet{zeiler2014visualizing} and \cite{dieleman2015rotation}
also use large receptive fields of $7\times7$ and $6\times6$
in the first convolution layer, respectively.
However, we find that using a receptive field larger than $5\times5$ in the first
convolutional layer leads to worse performance.
This result is in agreement with
\changed{
the network of \citet{simonyan2014very}, which has become known as ``VGGNet".
}
VGGNet features an extremely homogeneous architecture that only performs
$3\times3$ convolutions.
Using a large receptive field instead of a stack of multiple $3\times3$ convolutions
leads to a shallower network, and it is often preferable to increase
the depth by using smaller receptive fields.
However, we find that replacing the first layer with
a stack of two $3\times3$ convolutional layers increases the validation error,
\changed{
and thus use a $5\times5$ convolution in the first layer.
}

\changed{
In the remaining layers, we still follow VGGNet and add many
$3\times3$ convolutions (with zero-padding of size 1 pixel).
Note that with the padding of 1 pixel for $3\times3$ convolutional layers,
the spatial resolution will be preserved after convolution.
Such preservation of spatial resolution allows us to build relatively deep
networks, as shown in Table~\ref{table:hyperparamters}.
The main contribution of VGGNet is in showing that the depth plays an important
role in good performance.
In our case, we start with four convolutional layers and progressively add more
layers, while monitoring the validation loss;
we stop at eight convolutoinal layers after we find that adding more layers
leads to worse performance.
We conjecture that a greater depth and hence larger number of parameters
lead to overfitting in our case.
}

@@

\item{Reflection:
We flip each image horizontally with a probability of 0.5 to exploit mirror symmetry. }
\item{Translation:
We also have translational symmetry in the images.
Given an image of size $48\times48$ pixels, we extract a random contiguous crop
of size $44\times44$.
Each cropping is equivalent to randomly shifting a $44\times44$ image by up to 4 pixels
vertically and/or horizontally. }
\item{Gaussian noise:
We introduce random Gaussian noise to each pixel values
by using a similar method to \cite{krizhevsky2012imagenet}.}
\end{itemize}
\changed{
In addition to artificially increasing the size of the data set,
these data augmentation schemes make the resulting model more invariant to
rotation, reflection, translation, and small noise in the pixel values.
We also note that the data augmentation steps
}
add almost no computational cost,
as they are performed on the CPU while the GPU is training the ConvNets on images.

@@

To make final classifications,
we use our ConvNet model to make 64 sets of predictions for
64 transformations of the input images:
4 rotations, 4 horizontal translations, and 4 vertical translations
(with random horizontal reflections).
Although we use an identical network architecture for all transformations,
we consider each set of predictions as separate results from different models.
Finally, we use a model combination technique known as
Bayesian Model Combination~\citep[BMC;][]{Monteith2011}, which
uses Bayesian principles to generate an ensemble combination of different models.
Although the data augmentation step in Section~\ref{sec:data_augmentation} should
make our ConvNet model invariant to these types of transformations,
we find that applying BMC still results in a significant increase in performance.
\changed{
For a thorough discussion of BMC, we refer the reader to \citet{Monteith2011}
(See also \citet{carrascokind2014exhausting} for application of BMC to
photometric redshift estimation,
and \citet{kim2015hybrid} for combining star-galaxy classifiers).
}

\sout{ The posterior probability that a source is a galaxy is given by }
\sout{ $ P \left(G | \bmath{x}, \bmath{D}, \bmath{M}, \bmath{E} \right) = \sum_{e \in \bmath{E}} P \left(G | \bmath{x}, \bmath{M}, e \right) P \left(e | \bmath{D} \right), $ }
\sout{ where $\bmath{x}$ is the input, $\bmath{D}$ is the data set, $\bmath{M}$ is the set of models (\ie different transformations of input images), and $e$ is an element in the ensemble space $\bmath{E}$ of possible model combinations.  By Bayes' Theorem, the posterior probability of $e$ given $\bmath{D}$ is given by }
\sout{$ P \left(e | \bmath{D} \right) = \frac{P \left(e \right)}{P \left(\bmath{D} \right)} \prod_{d \in \bmath{D}} P \left( d | e \right) \propto P \left(e\right) \prod_{d \in \bmath{D}} P \left(d | e \right). $}
\sout{Here, $P\left(e \right)$ is the prior probability of $e$, which we assume to be uniform.  The product of $P\left(d | e\right)$ is over all individual data $d$ in the training data $\bmath{D}$, and $P\left(\bmath{D}\right)$ is merely a normalization factor and not important.  For probabilistic classifiers, we can directly use the probabilistic predictions and write Equation~\ref{eq:p_ensemble} as}
\sout{ $ P \left( e | \bmath{D} \right) \propto P \left( e \right) \prod_{i=1}^{N} \hat{y}_i y_i + \left(1 - \hat{y}_i\right) \left(1 - y_i\right), $}

\sout{Although the space $\bmath{E}$ of potential model combinations is in principle infinite, we can produce a reasonable finite set of potential model combinations by using sampling techniques.  In our implementation, the weights of each combination of the base classifiers is obtained by sampling from a Dirichlet distribution.  We first set all alpha values of a Dirichlet distribution to unity.  We then sample this distribution $q$ times to obtain $q$ sets of weights.  For each combination, we assume a uniform prior and calculate $P\left(e|\bmath{D}\right)$ using Equation~\ref{eq:p_ensemble_prob}.  We select the combination with the highest $P\left(e|\bmath{D}\right)$, and update the alpha values by adding the weights of the most probable combination to the current alpha values.  The next $q$ sets of weights are drawn using the updated alpha values.}

\sout{We continue the sampling process until we reach a predefined number of combinations, and we finally use Equation~\ref{eq:posterior_bmc} to compute the posterior probability that a source is a star (or a galaxy).  In this paper, we use a $q$ value of three, and we consider a total of 1,000 model combinations.}

@@

To compare the performance of ConvNets with
machine learning algorithms that use standard photometric features,
we use a machine learning framework called
Trees for Probabilistic Classifications (TPC).
TPC is a parallel, supervised machine learning algorithm
that uses prediction trees and random forest 
techniques~\citep{breiman1984classification, breiman2001random}
to produce a star-galaxy classification.
\changed{
A complete description of TPC is beyond the scope of this paper, and
we refer the reader to
\cite{carrascokind2013tpz} and \cite{kim2015hybrid} for more details.
While other random forest implementations exist, we have chosen TPC, because
it is tailored specifically for astronomical use~\citep{carrascokind2013tpz};
it has been tested for astronomical use cases, including
photometric redshift estimation~\citep{sanchez2014photometric}
and star-galaxy classification~\citep{kim2015hybrid};
it adds an extra layer of randomness by perturbing the measured attributes by
their uncertainties;
and it uses parallelism to handle large data sets on distributed memory systems.
}

>>>>

Almost all star-galaxy classifiers published in the literature use
the reduced summary information available from astronomical catalogs.
Constructing catalogs requires careful engineering and considerable domain expertise
to transform the reduced, calibrated pixel values that comprise an image into suitable features,
such as magnitudes or shape information of an object.
In a branch of machine learning called \textit{deep learning}~\citep{lecun2015deep},
features are not designed by human experts;
they are learned directly from data by deep neural networks.
Deep learning methods learn multiple levels of features by
transforming the feature at one level into a more abstract feature at a higher
level.
For example, when an array of pixel values is used as input to a deep learning
method, the features in the first layer might represent locations and orientations
of edges.
The second layer could assemble particular arrangements of edges into more
complex shapes, and subsequent layers would detect objects as
combinations of low-level features.
These multiple layers of abstraction progressively amplify aspects of the input
that are important for classification tasks.
Deep learning has been applied successfully to galaxy morphological
classification in Sloan Digital Sky Survey~\citep[SDSS;][]{dieleman2015rotation}
and Cosmic Assembly Near-infrared Deep Extragalactic Legacy
Survey~\citep[CANDELS;][]{huertas2015catalog} and to photometric redshift
estimation~\citep{hoyle2015measuring},
but it has not yet been applied to the problem of source classification.

@@

To demonstrate the performance of our ConvNet model, we use photometric and
spectroscopic data sets with different characteristics and compositions.
In this section, we briefly describe these data sets and the image pre-processing
steps for retrieving cutout images.

@@

Using the CasJobs server, we randomly select a total of 65,000 sources, which are either stars or galaxies.
We exclude some bad photometric observations as follows.
We consider only objects
with no warning flags in the spectroscopic measurement (\texttt{zWarning = 0});
the half-light radius in the $r$ band is less than 30 arc seconds
as measured by the exponential and de Vaucouleurs light profiles;
the error on the spectroscopic redshift measurement is less than 0.1; and
the spectroscopic redshift is less than 2.

To create training images,
we obtain the image FITS files for SDSS fields containing these objects
in five photometric bands: $u$, $g$, $r$, $i$, and $z$.
We use the astrometry information in the FITS headers
in the \texttt{Montage}\footnote{http://montage.ipac.caltech.edu/} software
to align each image to the reference ($r$-band) image.
We then use \texttt{SExtractor} to find the pixel positions of
the 65,000 objects we have selected,
and to center each object in the cutout image.
Magnitudes in the SDSS photometric catalog are expressed as inverse hyperbolic
sine magnitudes~\citep[also known as luptitudes;][]{lupton1999modified},
and we follow the SDSS convention and convert all flux values to luptitudes.
Finally, in order to account for the effect of Galactic dust,
extinction corrections in magnitudes are applied
following \cite{schlegel1998maps}.
In the end, we have cutout images of size $48\times48$ pixels
with luptitude values in each pixel.
We note that we have experimented with increasing the pixel dimensions to
$60\times60$ and $72\times72$ pixels, but do not find noticeable
improvement in the performance of our model.

In the end, we have 17,344 stars and 47,656 galaxies available
for the training and testing processes.
The apparent magnitudes range from $10.7 < r < 23.1$,
and the galaxies in this sample have a mean redshift of $z \sim 0.36$.
We randomly split the objects into training, held-out validation,
and blind test sets of size 40,000, 10,000, and 15,000, respectively.
We note that cross-validation is often avoided in deep learning
in favor of hold-out validation,
since cross-validation is computationally expensive.
We also note that we perform a blind test, and the test set is not used in any
way to train or calibrate the algorithms.
The first two panels of Figure~\ref{fig:sdss_mag} show the number of objects
and the fraction of stars in the test set as functions of $r$-band magnitude.
Similarly, Figure~\ref{fig:sdss_g_r} shows the number of objects and the
fraction of stars in the test set as functions of $g-r$ color.
The normalized kernel density estimate distributions for the training
and validation sets are almost identical to those of the test set,
and they are nearly indistinguishable when overlapped.
We do not show the distributions for the training and validation sets in
Figures~\ref{fig:sdss_mag} and \ref{fig:sdss_g_r} to avoid cluttering the
plots.

@@

In the end, we have 8,545 stars and 57,843 galaxies available
for the training and testing processes.
The apparent magnitudes range from $13.9 < r < 25.6$,
and the galaxies in this sample have a mean redshift of $z \sim 0.59$.
We randomly split the objects into training, held-out validation, and
blind test sets of size 40,000, 10,000, and 13,278, respectively.
Figures~\ref{fig:clens_mag} and \ref{fig:clens_g_r} show the distribution
of objects in the test set as functions of $i$-band magnitude and $g-r$ color.
We do not show the distributions for the training and validation sets,
since the normalized kernel density estimate distributions for the training
and validation sets are almost identical to those of the test set.

@@

Neural networks have many hyperparameters, including those that specify the
network itself (\eg the size and non-linearity of each layer)
and those that specify how the network is trained
(\eg the mini-batch size or the learning rate).
Furthermore, the architecture of a neural network can have a
significant impact on its performance.
In this section, we provide a brief description of key hyperparameters in our
ConvNet model, and also present the network architecture.

@@

An artificial neuron in most artificial neural networks is represented
as a mathematical function that models a biological neural structure
\citep{aggarwal2014data}.
A schematic representation is shown in Figure~\ref{fig:neuron_a}.
Let $\bmath{x}=\left(x_1,x_2,\dots,x_n\right)$ be a vector of inputs to a given neuron,
$\bmath{w}=\left(w_1,w_2,\dots,w_n\right)$ be a vector of weights, and
$b$ be the bias.
Then, the output of the neuron is
\citep{rosenblatt1961principles}
\begin{equation}
  y = \sigma \left( \bmath{w} \cdot \bmath{x} + b \right),
  \label{eq:neuron_output}
\end{equation}
where $\sigma$ is the activation function (or \textit{non-linearity}).
The most popular non-linearity at present
is the rectified linear unit
\citep[ReLU;][]{nair2010rectified}, $\sigma(x)=\max(0, x)$.
ReLUs generally allow much faster
training of deep neural networks with many layers.
However, ReLU units can sometimes result in dead neurons whose output is always zero.
To mitigate this problem, we use leaky ReLUs
~\citep{maas2013rectifier}
that have a small, non-zero slope in the negative region,
\begin{equation}
  \sigma(x) =
    \begin{cases}
      x & \mbox{if } x \geq 0 \\
      0.01x & \mbox{if } x < 0.
    \end{cases}
\end{equation}

Many deep learning models use feedforward neural network architectures with multiple layers,
where each neuron in one layer is connected to the neurons of the subsequent
layer~\citep{lecun2015deep}.
A schematic representation is shown in Figure~\ref{fig:neuron_b}.
All layers except the input and output layers are conveniently called hidden layers.

We
find a set of weights and biases such that,
given $N$ samples, the output from the network
$\bmath{y}=\left(y_1, y_2, \dots, y_N \right)$ approximates the desired output
$\hat{\bmath{y}}=\left(\hat{y}_1, \hat{y}_2, \dots, \hat{y}_N \right)$
as closely as possible for all input
$\bmath{X}=\left(\bmath{x}_1,\bmath{x}_2,\dots,\bmath{x}_N\right)$.
We can formulate this as the minimization of a loss function
$L(\bmath{y},\hat{\bmath{y}})$
over the training data.
In this work, we use \textit{cross-entropy}
\citep[also called log loss;][]{murphy2012machine} as the loss function.
For binary classification, the cross-entropy per sample is given by
\begin{equation}
  L(y_j, \hat{y}_j) = -\hat{y}_j \log_2 y_j - (1 - \hat{y}_j) \log_2(1 - y_j),
  \label{eq:cross_entropy}
\end{equation}
where $\hat{y}_j$ is the actual truth value (\eg 0 or 1) of the $j$-th data, and
$y_j$ is the probability prediction made by the model.
We compute the loss function by taking the average of all cross-entropies in the sample.
Thus, the loss function becomes
\begin{equation}
  L(\bmath{y}, \hat{\bmath{y}})
= - \frac{1}{N} \sum_{j=1}^{N} \hat{y}_j  \log_2 y_j
    + (1 - \hat{y}_j)  \log_2 (1 - \hat y_j).
  \label{eq:cross_entropy_all}
\end{equation}

To find the weights $\bmath{w}$ and biases $\bmath{b}$ which minimize the loss,
we use a technique called \textit{gradient descent},
where we use the following rules to update the parameters in each layer $l$:
\begin{align}
  \bmath{w}_l &\rightarrow
  \bmath{w}_l^{\prime}
  = \bmath{w}_l - \eta \frac{\partial L}{\partial \bmath{w}_l} \nonumber \\
  \bmath{b}_l &\rightarrow
  \bmath{b}_l^{\prime}
  = \bmath{b}_l - \eta \frac{\partial L}{\partial \bmath{b}_l},
  \label{eq:gradient_descent}
\end{align}
where $\eta$ is a small, positive number known as the \textit{learning rate}.
The gradients can be computed using the backpropagation
procedure~\citep{rumelhart1988learning}.
A common approach to speed up training is to split the training data
into mini-batches~\citep{lecun1998efficient}.
In mini-batch gradient descent, instead of computing the gradients in
Equation~\ref{eq:gradient_descent} for the entire training data,
we only compute the gradient of randomly chosen training examples at each step.
As training examples are usually correlated,
the gradient computed from each mini-batch is a good approximation
of the overall gradient~\citep{bottou1998online}.
As a result, mini-batch gradient descent results in much faster convergence.
However, there is a trade-off: the lower the batch size is,
the lower the convergence rate will be;
the higher the batch size is, the longer it will take to compute the gradient
at each step~\citep{bousquet2008tradeoffs}.
Thus, a moderate batch size, combined with a decaying learning rate, is
generally used in practice. We use a batch size of 128 in this work.

@@

@@

The convolutional neural network~\citep[ConvNet;][]{fukushima1980neocognitron,lecun1998gradient}
is a type of deep, feedforward neural network
that has recently become
a popular
approach in the computer vision community.
In a typical ConvNet, the first few stages are composed of two types of layers: 
convolutional layers and pooling layers.

The input to a convolutional layer is an image, and the output channels of
each layer are called \textit{feature maps}.
To produce output feature maps, we convolve each feature map with a set of weights
called \textit{filters},
and apply a non-linearity such as ReLU to the weighted sum of these convolutions.
Different feature maps use different sets of filters,
but all neurons in a feature map share the same set of filters.
Mathematically, we replace the dot product in Equation~\ref{eq:neuron_output}
with a sum of convolutions. Thus, the $k$-th feature map is given by
\begin{equation}
  y^k = \sigma \left( \sum_{m} \bmath{w}_m^k \ast \bmath{x}_m + b^k \right),
\end{equation}
where we sum over the set of input feature maps,
$\ast$ is the convolution operator, and $\bmath{w}_m^k$ represent the filters.

Typically, a
pooling layer computes the maximum of a local $2\times2$ patch in a feature map
\citep{krizhevsky2012imagenet}.
Since the pooling layer aggregates the activations of neighboring units from the previous layer,
it reduces the dimensionality of the feature maps and
makes the model invariant to small shifts and distortions
\citep{boureau2010theoretical}
.
Two or more layers of convolution and pooling are stacked,
followed by more convolutional and fully-connected layers.

@@

The architecture of \cite{krizhevsky2012imagenet}
uses relatively large receptive fields ($11\times11$) in the first convolutional layers.
\citet{zeiler2014visualizing} and \cite{dieleman2015rotation}
also use large receptive fields of $7\times7$ and $6\times6$
in the first convolution layer, respectively.
However, we find that using a receptive field larger than $5\times5$ in the first
convolutional layer leads to worse performance.
This result is in agreement with
the network of \citet{simonyan2014very}, which has become known as ``VGGNet".
VGGNet features an extremely homogeneous architecture that only performs
$3\times3$ convolutions.
Using a large receptive field instead of a stack of multiple $3\times3$ convolutions
leads to a shallower network, and it is often preferable to increase
the depth by using smaller receptive fields.
However, we find that replacing the first layer with
a stack of two $3\times3$ convolutional layers increases the validation error,
and thus use a $5\times5$ convolution in the first layer.

In the remaining layers, we still follow VGGNet and add many
$3\times3$ convolutions (with zero-padding of size 1 pixel).
Note that with the padding of 1 pixel for $3\times3$ convolutional layers,
the spatial resolution will be preserved after convolution.
Such preservation of spatial resolution allows us to build relatively deep
networks, as shown in Table~\ref{table:hyperparamters}.
The main contribution of VGGNet is in showing that the depth plays an important
role in good performance.
In our case, we start with four convolutional layers and progressively add more
layers, while monitoring the validation loss;
we stop at eight convolutoinal layers after we find that adding more layers
leads to worse performance.
We conjecture that a greater depth and hence larger number of parameters
lead to overfitting in our case.

@@

\item{Reflection:
We flip each image horizontally with a probability of 0.5 to exploit mirror symmetry. }
\item{Translation:
We also have translational symmetry in the images.
Given an image of size $48\times48$ pixels, we extract a random contiguous crop
of size $44\times44$.
Each cropping is equivalent to randomly shifting a $44\times44$ image by up to 4 pixels
vertically and/or horizontally. }
\item{Gaussian noise:
We introduce random Gaussian noise to each pixel values
by using a similar method to \cite{krizhevsky2012imagenet}.}
\end{itemize}
In addition to artificially increasing the size of the data set,
these data augmentation schemes make the resulting model more invariant to
rotation, reflection, translation, and small noise in the pixel values.
We also note that the data augmentation steps
add almost no computational cost,
as they are performed on the CPU while the GPU is training the ConvNets on images.

@@

To make final classifications,
we use our ConvNet model to make 64 sets of predictions for
64 transformations of the input images:
4 rotations, 4 horizontal translations, and 4 vertical translations
(with random horizontal reflections).
Although we use an identical network architecture for all transformations,
we consider each set of predictions as separate results from different models.
Finally, we use a model combination technique known as
Bayesian Model Combination~\citep[BMC;][]{Monteith2011}, which
uses Bayesian principles to generate an ensemble combination of different models.
Although the data augmentation step in Section~\ref{sec:data_augmentation} should
make our ConvNet model invariant to these types of transformations,
we find that applying BMC still results in a significant increase in performance.
For a thorough discussion of BMC, we refer the reader to \citet{Monteith2011}
(See also \citet{carrascokind2014exhausting} for application of BMC to
photometric redshift estimation,
and \citet{kim2015hybrid} for combining star-galaxy classifiers).

@@

To compare the performance of ConvNets with
machine learning algorithms that use standard photometric features,
we use a machine learning framework called
Trees for Probabilistic Classifications (TPC).
TPC is a parallel, supervised machine learning algorithm
that uses prediction trees and random forest 
techniques~\citep{breiman1984classification, breiman2001random}
to produce a star-galaxy classification.
A complete description of TPC is beyond the scope of this paper, and
we refer the reader to
\cite{carrascokind2013tpz} and \cite{kim2015hybrid} for more details.
While other random forest implementations exist, we have chosen TPC, because
it is tailored specifically for astronomical use~\citep{carrascokind2013tpz};
it has been tested for astronomical use cases, including
photometric redshift estimation~\citep{sanchez2014photometric}
and star-galaxy classification~\citep{kim2015hybrid};
it adds an extra layer of randomness by perturbing the measured attributes by
their uncertainties;
and it uses parallelism to handle large data sets on distributed memory systems.
